---
title: "Shellby Johnson- Lab 1"
author: "Shellby Johnson"
date: "January 9, 2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}

library(tidyverse)
library(here)
library(janitor)
library(kableExtra)

```

Read in noaa_fisheries.csv data
Use the "here" package which tells R where to find the file you need to read in

```{r}

us_landings <- read_csv(here("data", "noaa_fisheries.csv"))

```

Now, make a tidy version of the data frame:
- making all column names lower snake case
- clean names makes lower snake case in the janitor package
-str_to_lower makes the values in the columns specified lowercase

```{r}

landings_tidy <- us_landings %>% 
  janitor::clean_names() %>% 
  mutate(state = str_to_lower(state),
         afs_name = str_to_lower(afs_name)) %>% 
  mutate(dollars_num = parse_number(dollars_usd))


```


Get some info for salmon:

1st, remove the word "aggregate"
Then filter for everything in afs_clean that includes the word salmon

filter(str_detect(afs_clean, pattern = "salmon")) >>> does this pattern exists in this column, if so then "TRUE" and R returns only those

separate(afs_clean, into = c("group", "species"), sep = ",") >>> separate this column into the 2 columns "group" and "species" and you will separate it by the "," in afs_clean
```{r}

salmon_landings <- landings_tidy %>%
  mutate(afs_clean = str_remove(afs_name, pattern = "aggregate")) %>% 
  filter(str_detect(afs_clean, pattern = "salmon")) %>% 
  separate(afs_clean, into = c("group", "species"), sep = ",")

```


Find some group summary data

Find annual total US landings and dollar value (summing across all states) for each TYPE of salmon using 'group_by' +'summarize'

```{r}

salmon_summary <- salmon_landings %>% 
  group_by(year,species) %>% 
  summarize(
    tot_landings = sum(landings_pounds), 
    tot_value = sum(dollars_num)
  )

```


Make a grapgh!

```{r}

salmon_landings_graph <- ggplot(data = salmon_summary, aes(x = year, y = tot_landings)) +
  geom_line(aes(color = species))+
  theme_minimal()

salmon_landings_graph

```

Now we will export the graph with ggsave
ggsave will export a grapgh of your choosing, if will auto save the last graph you ran unless you specify

```{r}

ggsave(plot = salmon_landings_graph, 
       here("figures", "us_salmon_sj.png"),
       height = 5,
       width = 8)

```



Make a nice kable table
- only want 1st 5 lines

```{r}

salmon_5 <- salmon_summary %>% 
  head(5)

```

Use Kable

```{r}

kable(salmon_5) %>% 
  kable_styling(bootstrap_options = "striped",
                full_width = FALSE)


```

